Fast and adaptive cointegration based model for forecasting high frequency financial time series

### 2017-01-13 articles-conferences/iscef2016/body.tex
Review Intro
General comments 3 ready

Our aims were to study if a parallel version of VECM can be used with high
frequency stream data.  Our approach was to determine adaptively the number of
observations and lags of VECM which maximise cointegration relations in the
past.  This search is computationally expensive because the Johansen method, see
\cite{johansen1995}, is required to find cointegration vectors, which is of
order $O(n^3)$. Therefore, our proposal, called Adaptively Vector Error
Correction (AVECM), is to parallelise this search in order to get new parameters
before new data arrives.  Model effectiveness is focused on out-of-sample
forecast rather than in-sample fitting. This criterion allows AVECM prediction
capability to be expressed rather than just explaining data history.  The
forecast capability of our method was measured using MSE and the Theil's
$U$-statistic, see \cite{theil1966}, widely used in economic forecast. Tests
were run using four currency rates: Euro (EUR) to United States Dollar (USD)
(EURUSD), British Pound (GBP) to USD (GBPUSD), USD to Swiss Franc (CHF) (USDCHF)
and USD to Japanese Yen (JPY) (USDJPY) with a 10-second frequency.

>>>>

Cointegration relationship can be found in low and high frequency data of two
or more assets. While cointegration in low frequency data is motivated by a
long-run equilibrium relationship between economic forces, cointegration in
high frequency data has its foundation in statistical arbitrage which is very
helpful to detect mean-reverting trades. Cointegration at different data
frequencies may differ. \cite{rittler2012} shows how cointegration appears when
the data frequency is increased. Our aims were to study if a parallel
version of VECM can be used with high frequency stream data.

Our approach was to determine adaptively the number of observations and lags of
VECM which maximise cointegration relations in the past.  This search is
computationally expensive because the Johansen method, see \cite{johansen1995},
is required to find cointegration vectors, which is of order $O(n^3)$.
Therefore, our proposal, called Adaptively Vector Error Correction (AVECM), is
to parallelise this search in order to get new parameters before new data
arrives.  Model effectiveness is focused on out-of-sample forecast rather than
in-sample fitting. This criterion allows AVECM prediction capability to be
expressed rather than just explaining data history.  The forecast capability of
our method was measured using MSE and the Theil's $U$-statistic, see
\cite{theil1966}, widely used in economic forecast. Tests were run using four
currency rates: Euro (EUR) to United States Dollar (USD) (EURUSD), British
Pound (GBP) to USD (GBPUSD), USD to Swiss Franc (CHF) (USDCHF) and USD to
Japanese Yen (JPY) (USDJPY) with a 10-second frequency.
